A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2022-11-10 analysis/paper/paper.Rmd
Set up as.character(english::as.english()) to print numbers in written form a couple of places

```{r libraries, eval = TRUE, results = 'hide'}
library(here)
library(tidyverse)
library(sf)
library(ggridges)
library(ggthemes)
library(ggnewscale)
library(patchwork)
library(terra)
library(raster)
library(oxcAAR)
library(vapour)
library(topoDistance)
library(IRanges)
library(stringr)
library(kableExtra)
library(english)
quickSetupOxcal()

@@

Another immediately striking result is the apparent deviation from the shoreline towards the end of the Stone Age. After around 2500 BCE several sites are situated a considerable distance from the reconstructed shoreline, and while a couple remain horizontally and topographically close, most appear to be considerably elevated above the sea-level. Of the results from after 2500 BCE, which are associated with `r length(unique(r2$sitename))` sites, only `r as.english(r2a0)` has simulation results for vertical distance that includes zero. The highest simulated vertical distance among these is `r round(rs2$max, 1)`m and the median is `r round(rs2$median, 1)`m. Furthermore, some deviation is evident from just after 4000 BCE we well. The sample size is limited, but of the `r length(unique(r3$sitename))` sites associated with the period between 4000 and 2500 BCE, `r as.english(r3a25)` sites have all vertical distance results above 25m. However, the median vertical distance of the results from this period is only `r round(rs3$median, 1)`m, indicating that while some sites have a markedly withdrawn location, most are still situated close to the shoreline. The chronological smearing following from the uncertainty in the ^14^C-dates means that while the results cannot be used to directly inform discussions that deal with the century scale around these chronological transitions [e.g.  @prescott2020; @solheim2021], the findings are nonetheless in clear agreement with the general chronological developments suggested in the literature.

>>>>

```{r libraries, eval = TRUE, results = 'hide'}
library(here)
library(tidyverse)
library(sf)
library(ggridges)
library(ggthemes)
library(ggnewscale)
library(patchwork)
library(terra)
library(raster)
library(oxcAAR)
library(topoDistance)
library(IRanges)
library(stringr)
library(kableExtra)
library(english)
quickSetupOxcal()

@@

Another immediately striking result is the apparent deviation from the shoreline towards the end of the Stone Age. Of the results from after 2500 BCE, which are associated with `r length(unique(r2$sitename))` sites, only `r as.character(as.english(r2a0))` has simulation results for vertical distance that includes zero. The highest simulated vertical distance among these is `r round(rs2$max, 1)`m and the median is `r round(rs2$median, 1)`m. Furthermore, some deviation from the shoreline is evident from just after 4000 BCE as well. The sample size is limited, but of the `r length(unique(r3$sitename))` sites associated with the period between 4000 and 2500 BCE, `r as.character(as.english(r3a25))` sites have all vertical distance results above 25m. However, the median vertical distance of the results from this period is only `r round(rs3$median, 1)`m, indicating that while some sites have a markedly withdrawn location, most are still situated close to the shoreline. The chronological smearing following from the uncertainty in the ^14^C-dates means that while the results cannot be used to directly inform discussions that deal with the century scale around these chronological transitions [e.g.  @prescott2020; @solheim2021], the findings are nonetheless in clear agreement with the general chronological developments suggested in the literature.
